---
title: "First steps - Install and Run AlphaSimHlpR tutorial"
author: "Marnin Wolfe"
date: "2021-04-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preamble: Install AlphaSimHlpR

R packages we will need. Install them if necessary.

```{r, eval=F}
install.packages(c("tidyverse","AlphaSimR","devtools"))
```

Install AlphaSimHlpR

```{r, eval=F}
devtools::install_github("jeanlucj/AlphaSimHlpR", ref = 'master', 
                         dependencies = T, force = T) # force = T to ensure I get a fresh install
```

When prompted "Which would you like to update?" choose "1: All".

```{r, eval=F}
library(AlphaSimHlpR)
# Get `Error: package ‘optiSel’ could not be loaded`??
install.packages("optiSel",dependencies = T)
library(AlphaSimHlpR)
# still error
library(optiSel)
# Error: package or namespace 
# load failed for ‘optiSel’: .on
# Load failed in loadNamespace() for 'rgl', 
# details: call:         rgl.init(initValue, onlyNULL) 
# error: OpenGL is not available in this build
```

**Following is specific to my macOS install state**

Google search of error leads to: <https://stackoverflow.com/questions/9878693/error-in-loading-rgl-package-with-mac-os-x>

Suggestion Solution: install XQuartz

```{bash, eval=F}
brew install xquartz
```

```{r, eval=F}
library(AlphaSimHlpR)
```

Finally I get a clean load!

```{r, eval=F}
browseVignettes("AlphaSimHlpR")
```

The vignettes don't show up... but their Rmd's are in the GitHub Repo. Best guess: need to be added to the `namespace` or `knit` and the package master needs to be freshly built.

I downloaded the Rmd's [from GitHub here](https://github.com/jeanlucj/AlphaSimHlpR/tree/master/vignettes).

# Basic AlphaSimHlpR tutorial

New R session. Follow the AlphaSimHlpR vignette.

**I also had to download the `inst` folder and it's example "control file" contents [from GitHub here](https://github.com/jeanlucj/AlphaSimHlpR/tree/master/inst)**

```{r}
# Make sure you have the right packages installed
neededPackages <- c("AlphaSimR", "dplyr", "tidyr", "plotrix", 
                    "lme4", "sommer", "optiSel")
for (p in neededPackages) if (!require(p, character.only=T)) install.packages(p)
suppressMessages(library(AlphaSimHlpR))
```

## Define simulation settings

Define the genetic architecture of the population and other breeding scheme parameters in a list `bsp`.

```{r}
bsp <- specifyPopulation(ctrlFileName="data/inst/PopulationCtrlFile_Small.txt")
bsp <- specifyPipeline(bsp, ctrlFileName="data/inst/ControlFile_Small.txt")
bsp <- specifyCosts(bsp, ctrlFileName="data/inst/CostsCtrlFile_Small.txt")
nReplications <- 3
bsp$nCyclesToRun <- 6

print(bsp)
```

## Replicate simple scheme

Run a simple breeding scheme for `r bsp$nCyclesToRun` cycles

Replicate a very simple breeding program `r nReplications` times.

```{r}
replicRecords <- lapply(1:nReplications, runBreedingScheme, 
                        nCycles=bsp$nCyclesToRun, 
                        initializeFunc=initFuncADChk, 
                        productPipeline=prodPipeFncChk, 
                        populationImprovement=popImprov1Cyc, bsp)
```

## Calculate means and plot results

Calculate the means of the breeding programs and plot them out

```{r}
plotData <- plotRecords(replicRecords)
meanMeans <- tapply(plotData$genValMean, list(plotData$year, plotData$stage), mean)
meanMeans <- meanMeans[,c("F1", bsp$stageNames)]
stdErrMeans <- tapply(plotData$genValMean, list(plotData$year, plotData$stage), std.error)
stdErrMeans <- stdErrMeans[,c("F1", bsp$stageNames)]
print(meanMeans)
print(stdErrMeans)
```

# Next step

Run a simple example [simulation of the effect of reducing error with new tools](example-simulation-reducing-error-with-new-tools.html).

