---
title: "Example simulation - reducing error with new tools"
author: "Marnin Wolfe"
date: "2021-04-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Previously

[AlphaSimHlpR first steps](AlphaSimHlpR_firstSteps.html): Installation and quickly running the `AlphaSimHlpR` tutorial example.

# Simulation with a burn-in period

Generate a baseline population. As this is an example, I won't use empirical inputs. Ignore costs for now.

-   Initialized with the provided "small" example control files. Made some minor manual changes, which I will print below.
-   See the files `BaselineControlFile.txt` and `BaselinePopulationFile.txt` in this repository.

I want to set-up "burn-in" generations of phenotypic selection. See if I can work that in `AlphaSimHlpR`.

**SPOILER ALERT: Not quite working yet. More work needed.**

The first step is to read in the control files defining the baseline breeding program and other breeding scheme parameters in a list `bsp`.

The tutorial example repeatedly runs the `runBreedingScheme()` function, which in turn depends on specified functions: `initializeFunc=initFuncADChk, productPipeline=prodPipeFncChk, populationImprovement=popImprov1Cyc`.

`selCritPipeAdv` and `selCritPopImprov` are set to **selCritIID**. 

Now, to continue, we can't use the `runBreedingScheme()` function because it will re-run the `initializeFunc()` and subsequently `runMacs2()`, which overwrites our founder haplotypes.

Make my own function: `runBreedingScheme_wBurnIn()`. 

Specify number of burn-in and post-burn in cycles. 

Also let's uses two separate `bsp` objects, one for before, another for after. *Use with caution.*

```{r}
rm(list=ls()); gc()
library(tidyverse); library(magrittr); 
suppressMessages(library(AlphaSimHlpR))
bsp <- specifyPopulation(ctrlFileName="data/BaselinePopulationFile.txt")
bsp <- specifyPipeline(bsp, ctrlFileName="data/BaselineControlFile.txt")
```

```{r}
runBreedingScheme_wBurnIn <- function(replication=NULL, 
                                      nBurnInCycles=10,nPostBurnInCycles=10,
                                      bspBurnIn,bspPostBurnIn,
                                      initializeFunc, 
                                      productPipeline, 
                                      populationImprovement){
  
  cat("******", replication, "\n")
  
  # This initiates the founding population 
  initList <- initializeFunc(bspBurnIn)
  SP <- initList$SP
  bspBurnIn <- initList$bsp
  records <- initList$records
  
  # Burn-in cycles
  for (cycle in 1:nBurnInCycles){
    cat(cycle, " ")
    records <- productPipeline(records, bspBurnIn, SP)
    records <- populationImprovement(records, bspBurnIn, SP)
  }
  # records <- AlphaSimHlpR:::lastCycStgOut(records, bspPostBurnIn, SP)

  cat("\n")
  cat("Burn-in Cycles")
  cat("\n")
  # Post burn-in cycles
  
  for (cycle in (nBurnInCycles+1):(nBurnInCycles+nPostBurnInCycles)){
    cat(cycle, " ")
    records <- productPipeline(records, bspPostBurnIn, SP)
    records <- populationImprovement(records, bspPostBurnIn, SP)
  }
  cat("\n")
  # Finalize the stageOutputs
  records <- AlphaSimHlpR:::lastCycStgOut(records, bspPostBurnIn, SP)
  
  return(list(records=records,
              bspBurnIn=bspBurnIn,
              bspPostBurnIn=bspPostBurnIn,
              SP=SP))
}
```

For burn-in, using phenotypic selection, so the `bsp` already created.

For post burn-in, change `selCritPopImprov` to `selCritGRM`, but ignore `selCritPipeAdv` for now.

```{r, eval=F}
bspBurnIn<-bsp
bspPostBurnIn<-bspBurnIn
bspPostBurnIn[["selCritPopImprov"]] <- selCritGRM

test_sim<-runBreedingScheme_wBurnIn(replication = 1, 
                                    nBurnInCycles=2,nPostBurnInCycles=2,
                                    bspBurnIn=bspBurnIn,
                                    bspPostBurnIn=bspPostBurnIn,
                                    initializeFunc=initializeScheme, 
                                    productPipeline=productPipeline, 
                                    populationImprovement=popImprov1Cyc)
```

Unfortunately, haven't gotten this to work yet. It seems likely more tweaking to the package is necessary.

I traced the problem at least as far as the `populationImprovement()` function in the post-burn-in phase and the `makeGRM()`.

```{r, eval=F, echo=F}
### DEBUGGING
# bspBurnIn<-bsp
# bspPostBurnIn<-bspBurnIn
# bspPostBurnIn[["selCritPopImprov"]] <- selCritGRM
# replication = 1; 
# nBurnInCycles=2;nPostBurnInCycles=2;
# bspBurnIn=bspBurnIn;
# bspPostBurnIn=bspPostBurnIn;
# initializeFunc=initializeScheme; 
# productPipeline=productPipeline; 
# populationImprovement=popImprov1Cyc
```

# Simulation - no burn-in

Set-up an experiment. No burn-in simulations. 
```{r}
# rm(list=ls()); gc()
# library(tidyverse); library(magrittr); 
# suppressMessages(library(AlphaSimHlpR))
# bsp <- specifyPopulation(ctrlFileName="BaselinePopulationFile.txt")
# bsp <- specifyPipeline(bsp, ctrlFileName="BaselineControlFile.txt")
```

```{r}
# test<-runBreedingScheme(replication = 1, nCycles = 5, # just 5 for now 
#                         initializeFunc=initializeScheme, 
#                         productPipeline=prodPipeFncChk, 
#                         populationImprovement=popImprov1Cyc, bsp)
# 
# test_burnin$records$stageOutputs %>% ggplot(.,aes(x=year,y=genValMean, color=stage)) + geom_line()
# test_burnin$records
```
